### 2020-02-17 Paper.tex
Upload Files

% !TEX program = xelatex
\documentclass[10pt,journal,final,a4paper,nofonttune]{IEEEtran}
\usepackage[OT1]{fontenc}

\title{Demo for IEEE Journals}
\author{Yifan Yang, Liye Jia, Kai Lin, Yujie Wang}

\begin{document}
\maketitle

\begin{abstract}
    Fuck this document! This is supposed to be an abstract no shorter 
    than 100 words. However, I cannot do that myself.
\end{abstract}

\begin{IEEEkeywords}
    Fuck, Documents, No Keywords
\end{IEEEkeywords}

% \tableofcontents
% \pagebreak

\section{Introduction}
This is going to be an essay that follows the IEEE standard.
I am now writing casually without thinking what I am doing. So, the 
sentence may seem irreadable.
\cite{gomez2012overview}

This is another paragraph in the Introduction Section

\section{Application Domain}
Paragraph one Paragraph one Paragraph one Paragraph one Paragraph one 
Paragraph one Paragraph one Paragraph one Paragraph one Paragraph one 
Paragraph one Paragraph one Paragraph one Paragraph one Paragraph one 
Paragraph one Paragraph one Paragraph one Paragraph one Paragraph one 
Paragraph one Paragraph one Paragraph one Paragraph one Paragraph one

Paragraph two Paragraph two Paragraph two Paragraph two Paragraph two 
Paragraph two Paragraph two Paragraph two Paragraph two Paragraph two 
Paragraph two Paragraph two Paragraph two Paragraph two Paragraph two 
Paragraph two Paragraph two Paragraph two Paragraph two Paragraph two 
Paragraph two Paragraph two Paragraph two Paragraph two Paragraph two

\section{Use Case Scenarios}
Paragraph one Paragraph one Paragraph one Paragraph one Paragraph one 
Paragraph one Paragraph one Paragraph one Paragraph one Paragraph one 
Paragraph one Paragraph one Paragraph one Paragraph one Paragraph one 
Paragraph one Paragraph one Paragraph one Paragraph one Paragraph one 
Paragraph one Paragraph one Paragraph one Paragraph one Paragraph one

Paragraph two Paragraph two Paragraph two Paragraph two Paragraph two 
Paragraph two Paragraph two Paragraph two Paragraph two Paragraph two 
Paragraph two Paragraph two Paragraph two Paragraph two Paragraph two 
Paragraph two Paragraph two Paragraph two Paragraph two Paragraph two 
Paragraph two Paragraph two Paragraph two Paragraph two Paragraph two

\section{System Description}

Paragraph one Paragraph one Paragraph one Paragraph one Paragraph one 
Paragraph one Paragraph one Paragraph one Paragraph one Paragraph one 
Paragraph one Paragraph one Paragraph one Paragraph one Paragraph one 
Paragraph one Paragraph one Paragraph one Paragraph one Paragraph one 
Paragraph one Paragraph one Paragraph one Paragraph one Paragraph one 
Here is an equation:

Paragraph two Paragraph two Paragraph two Paragraph two Paragraph two 
Paragraph two Paragraph two Paragraph two Paragraph two Paragraph two 
Paragraph two Paragraph two Paragraph two Paragraph two Paragraph two 
Paragraph two Paragraph two Paragraph two Paragraph two Paragraph two 
Paragraph two Paragraph two Paragraph two Paragraph two Paragraph two

\section{Conclusion}

Paragraph one Paragraph one Paragraph one Paragraph one Paragraph one 
Paragraph one Paragraph one Paragraph one Paragraph one Paragraph one 
Paragraph one Paragraph one Paragraph one Paragraph one Paragraph one 
Paragraph one Paragraph one Paragraph one Paragraph one Paragraph one 
Paragraph one Paragraph one Paragraph one Paragraph one Paragraph one

Paragraph two Paragraph two Paragraph two Paragraph two Paragraph two 
Paragraph two Paragraph two Paragraph two Paragraph two Paragraph two 
Paragraph two Paragraph two Paragraph two Paragraph two Paragraph two 
Paragraph two Paragraph two Paragraph two Paragraph two Paragraph two 
Paragraph two Paragraph two Paragraph two Paragraph two Paragraph two

\section*{Acknowledgments}

Paragraph one Paragraph one Paragraph one Paragraph one Paragraph one 
Paragraph one Paragraph one Paragraph one Paragraph one Paragraph one 
Paragraph one Paragraph one Paragraph one Paragraph one Paragraph one 
Paragraph one Paragraph one Paragraph one Paragraph one Paragraph one 
Paragraph one Paragraph one Paragraph one Paragraph one Paragraph one

\bibliographystyle{IEEEtran}
\bibliography{References}
\end{document}